---
output:
  md_document:
    variant: gfm
html_preview: false
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

# ggtreeExtra: Compact visualization of richly annotated phylogenetic data

If you use this work in published research, please cite:

S Xu, Z Dai, P Guo, X Fu, S Liu, L Zhou, W Tang, T Feng, M Chen, L Zhan, T Wu, E Hu, Y Jiang^\*^, X Bo^\*^ and __G Yu__<sup>\*</sup>. ggtreeExtra: Compact visualization of richly annotated phylogenetic data. __*Methods in Ecology and Evolution*__. 2021, msab166. doi: [10.1093/molbev/msab166](https://doi.org/10.1093/molbev/msab166) (accepted).


This repo contains source code and data to produce <!--Manuscript and-->Supplementary Material of the above paper.

  + rawdata: contains `HMP_tree`, `kegg` and `PhyloPhlAn`, downloaded from the examples of [GraPhlAn](https://github.com/biobakery/graphlan/tree/master/examples), and the `Methanotroph`, downloaded from the [repo](https://github.com/TheWrightonLab/Methanotroph_rpS3Analyses_SmithWrighton2018).
  + scripts: contains the script to produce the dataset of `data` using the data set of `rawdata`.
  + data: contains all the data sets that used to generate figures in Supplemental file.
  + Rmarkdown: contains the source code to produce Supplementary File.


## Dependencies and locations
  <!-- - GNU Make should be located in the user’s PATH -->
  <!-- - Python (2 or 3) should be installed and located in the user’s PATH -->
  <!--and `Biopython` also should be installed. -->

  + R (>= 4.0.3) should be installed and located in the user's PATH, and the following packages should be installed.
    - `ggtreeExtra` :                                          `BiocManager::install("ggtreeExtra")`
    - `ggtree` :                                               `BiocManager::install("ggtree")`
    - `treeio` :                                               `BiocManager::install("treeio")`
    - `tidytree` :                                             `install.packages("tidytree")`
    - `ggplot2` :                                              `install.packages("ggplot2")`
    - `kableExtra` :                                           `install.packages("kableExtra")`
    - `bookdown` :                                             `install.package("bookdown")`
    - `MicrobiotaProcess` :                                    `BiocManager::install("MicrobiotaProcess")`
    - `ggstar` :                                               `install.packages("ggstar")`
    - `Cairo` :                                                `install.packages("Cairo")`
    - `aplot` :                                                `install.packages("aplot")`
    - `patchwork` :                                            `install.packages("patchwork")`
    - `ggnewscale` :                                           `install.packages("ggnewscale")`
    - `knitr` :                                                `install.packages("knitr")`
    - `ggpmisc` :                                              `install.packages("ggpmisc")`
    - [`ggpattern`](https://github.com/coolbutuseless/ggpattern) :      `remotes::install_github("coolbutuseless/ggpattern")`
    - `tibble` :                                             `install.packages("tibble")`
    - `tidyr` :                                              `install.packages("tidyr")`
    - `dplyr` :                                              `install.packages("dplyr")`
    - `ggimage` :                                             `install.packages("ggimage")`
    - `ggridges` :                                            `install.packages("ggridges")`
  
```{r, message=FALSE, echo=FALSE, setup}
library(ggtreeExtra)
library(ggstar)
library(ggplot2)
library(ggtree)
library(treeio)
library(tidytree)
library(ggnewscale)
library(MicrobiotaProcess)
library(aplot)
library(kableExtra)
library(Cairo)
library(patchwork)
library(knitr)
library(ggpattern)
library(ggpmisc)
library(ggridges)
library(tidyr)
library(dplyr)
library(tibble)
library(ggimage)
```

To compile the <!-- Rmarkdown/manuscript.docx and --> Rmarkdown/[supplemental_file.pdf](Rmarkdown/supplementary_file.pdf), please run the following command on terminal.

<!-- ```r
make manuscript && make supple 
``` -->

```bash
Rscript -e 'rmarkdown::render("./Rmarkdown/supplementary_file.Rmd")'
```

Or run the following command in R.

```r
rmarkdown::render("./Rmarkdown/supplementary_file.Rmd")
```

Here is the output of `sessionInfo()` of the system on which [the Supplemental file](Rmarkdown/supplementary_file.pdf) was compiled:

```{r, echo=FALSE}
sessionInfo()
```

## Docker image

We also provided a [docker image](https://hub.docker.com/r/xushuangbin/ggtreeextraarticleenv) to help users to build the computing environment. You can pull and run it according to the following commands.

- Install Docker (https://www.docker.com/)
  - `sudo apt-get install docker.io` (Ubuntu)
- Pull the Docker image from Docker Hub:
  - `docker pull xushuangbin/ggtreeextraarticleenv:latest`
  - or
  - `sudo pull xushuangbin/ggtreeextraarticleenv:latest`
- Run the image:
  - `docker run -e PASSWORD=yourpassword -p 8787:8787 xushuangbin/ggtreeextraarticleenv`
  - or
  - `sudo docker run -e PASSWORD=yourpassword -p 8787:8787 xushuangbin/ggtreeextraarticleenv`
- Log in to RStudio at [http://localhost:8787](http://localhost:8787) using username `rstudio` and password `yourpassword`. For Windows users, you also need to provide your IP address, you can find it using `docker-machine ip default`.
+ Inside the RStudio, run:
  `browseVignettes(package = "ggtreeExtraArticleEnv")`
- You can click one of the links: "PDF", "source", "R code"
- In case of `The requested page was not found` error, try add 'help/' in front of the hostname in the URL (this is a known bug):
  http://localhost:8787/help/library/ggtreeExtraArticleEnv/doc/supplementary_file.pdf
