---
output:
  md_document:
    variant: gfm
html_preview: false
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

# ggtreeExtra: An R package to visualize compact circular layers of phylogenetic tree

<!--If you use this work in published research, please cite:
    Shuangbin Xu, , Zehan Dai, Pingfan Guo, Xiaocong Fu, Shanshan Liu, Lang Zhou, Wenli Tang, Tingze Feng, Meijun Chen, Li Zhan and GuangChuang Yu*. ggtreeExtra: An R package to visualize compact circular layers of phylogenetic tree, XXX, XXX, XXX -->

----

This repo contains source code and data to produce Manuscript and Supplementary Material of the above paper.

  + rawdata: contains `HMP_tree`, `kegg` and `PhyloPhlAn`, downloaded from the examples of [GraPhlAn](https://github.com/biobakery/graphlan/tree/master/examples), and the `Methanotroph`, downloaded from the [repo](https://github.com/TheWrightonLab/Methanotroph_rpS3Analyses_SmithWrighton2018).
  + data: contains the data has been processed, it can be displayed directly in the paper.
  + Rmarkdown: contains the source code to produce Manuscript and Supplementary Material.
  + scripts: contains the script to produce the dataset of `data` using the data set of `rawdata`.

## Dependencies and locations

  + GNU Make should be located in the user's PATH
  + Python (2 or 3) should be installed and located in the user's PATH 
    and `Biopython` also should be installed.
  + R (>=4.0.3) should be installed and located in the user's PATH
    + the following packages also should be installed.
      `ggtreeExtra`, `ggtree`, `tidytree`, `treeio`, `MicrobiotaProcess`, `kableExtra`, `ggplot2`, 
      `bookdown`, `ggstar`, `tidyverse`, `Cairo`, `openxlsx`, `patchwork`, `ggnewscale` and `knitr`.

```{r, message=FALSE, echo=FALSE, setup}
library(ggtreeExtra)
library(ggstar)
library(ggplot2)
library(ggtree)
library(treeio)
library(tidytree)
library(ggnewscale)
library(MicrobiotaProcess)
library(tidyverse)
library(openxlsx)
library(kableExtra)
library(Cairo)
library(patchwork)
library(knitr)
```

To compile the Rmarkdown/manuscript.docx and Rmarkdown/supplemental_file.pdf, please run the following command on terminal of Linux or Mac.

<!-- ```r
make manuscript && make supple 
``` -->

```r
make supple
```

Here is the output of `sessionInfo()` on the system on which [the document](https://github.com/xiangpin/ggtreeExtra_article/blob/master/Rmarkdown/supplementary_file.pdf) was compiled:

```{r, echo=FALSE}
sessionInfo()
```

