---
output:
  md_document:
    variant: gfm
html_preview: false
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

# ggtreeExtra: Compact visualization of richly annotated phylogenetic data

If you use this work in published research, please cite:

S Xu, Z Dai, P Guo, X Fu, S Liu, L Zhou, W Tang, T Feng, M Chen, L Zhan and __G Yu__<sup>\*</sup>. ggtreeExtra: Compact visualization of richly annotated phylogenetic data. _in preparation_.


This repo contains source code and data to produce <!--Manuscript and-->Supplementary Material of the above paper.

  + rawdata: contains `HMP_tree`, `kegg` and `PhyloPhlAn`, downloaded from the examples of [GraPhlAn](https://github.com/biobakery/graphlan/tree/master/examples), and the `Methanotroph`, downloaded from the [repo](https://github.com/TheWrightonLab/Methanotroph_rpS3Analyses_SmithWrighton2018).
  + scripts: contains the script to produce the dataset of `data` using the data set of `rawdata`.
  + data: contains all the data sets that used to generate figures in Supplemental file.
  + Rmarkdown: contains the source code to produce Supplementary File.


## Dependencies and locations
  <!-- - GNU Make should be located in the user’s PATH -->
  <!-- - Python (2 or 3) should be installed and located in the user’s PATH -->
  <!--and `Biopython` also should be installed. -->

  + R (>= 4.0.3) should be installed and located in the user's PATH, and the following packages should be installed.
    - `ggtreeExtra`
    - `ggtree`
    - `treeio`
    - `tidytree`
    - `ggplot2`
    - `kableExtra`
    - `bookdown`
    - `MicrobiotaProcess`
    - `ggstar`
    - `Cairo`
    - `aplot`
    - `patchwork`
    - `ggnewscale`
    - `knitr`
    - `ggpmisc`
    - [`ggpattern`](https://github.com/coolbutuseless/ggpattern)
    - `tibble`
    - `tidyr`
    - `dplyr`
    - `ggimage`
    - `ggridges`
  
```{r, message=FALSE, echo=FALSE, setup}
library(ggtreeExtra)
library(ggstar)
library(ggplot2)
library(ggtree)
library(treeio)
library(tidytree)
library(ggnewscale)
library(MicrobiotaProcess)
library(aplot)
library(kableExtra)
library(Cairo)
library(patchwork)
library(knitr)
library(ggpattern)
library(ggpmisc)
library(ggridges)
library(tidyr)
library(dplyr)
library(tibble)
library(ggimage)
```

To compile the <!-- Rmarkdown/manuscript.docx and --> Rmarkdown/[supplemental_file.pdf](Rmarkdown/supplemental_file.pdf), please run the following command on terminal.

<!-- ```r
make manuscript && make supple 
``` -->

```bash
Rscript -e 'rmarkdown::render("./Rmarkdown/supplementary_file.Rmd")'
```

Or run the following command in R.

```r
rmarkdown::render("./Rmarkdown/supplementary_file.Rmd")
```

Here is the output of `sessionInfo()` of the system on which [the Supplemental file](Rmarkdown/supplementary_file.pdf) was compiled:

```{r, echo=FALSE}
sessionInfo()
```

